Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-21 cav2017/cav2017.tex
Added todo notes package, also moved safety section
(I like todo notes)

\usetikzlibrary{positioning, automata, shapes.arrows, calc, shapes, arrows}
\usetikzlibrary{patterns}

\newcommand{\todo}[1]{{\color{red} TODO: #1}}
\newcommand{\commentib}[1]{{\color{blue} [IB: #1]}}
\newcommand{\mat}[1]{\boldsymbol{#1}}

@@

\begin{abstract}
This work presents a sound and automated approach to synthesise 
digital feedback control architectures for physical plants represented 
as linear, time invariant models. Models are encompassed by dynamical 
equations with inputs, evolving over a continuous state space describing 
the evolution of the physical variables of the system.  The approach is divided 
in two main parts, brought together by a counter-example guided inductive 
synthesis (CEGIS) loop, as follows. First, we devise a static feedback controller 
that stabilises the system; then, we verify its potential safety by means of either 
of the following approaches: forward reachability analysis via unfolding of the dynamics, 
or invariant set generation via abstract acceleration;  
further, if the above step fails, we loop back and invoke a CEGIS call 
to seek alternative stabilising feedbacks, benefitting from the obtained counter-examples 
to safety. 
\textcolor{red}{[Mention GA and Abstract Acceleration here?]}  
The approach is proven sound by accounting for errors due to the digitalisation of signals that are manipulated by the controller block.  
The proposed synthesis approaches and a developed tool are evaluated considering 
a set of intricate physical plant models taken from the digital control literature.
\keywords{
State-space dynamical models of physical systems; 
Digital controllers; 
Analogue-to-digital converters; 
Time sampling; 
Quantisation; 
Fixed-point arithmetics; 
CEGIS; 
safety requirements. 
}
\end{abstract}

@@

In this work, we propose the synthesis of \emph{safe} control algorithms for
LTI models taking into consideration the continuous domain of the plant, 
alongside the discrete domain of the controller and the hybrid elements that relate them.  
%% In this paper, we synthesize controllers for hybrid
%% closed-loop systems, i.e., software-implemented embedded controllers along
%% with a model of their physical environment (the plant), that are
%% \emph{safe}.
Due to the complexity of such systems, in this particular work we focus
on linear models \todo{[clarify next] with known configurations} and perform parametric synthesis
of safe digital controllers.

@@

\todo{Add contributions (Cristina)}

@@

%\todo{[Move material from \ref{sec:rw} here]}

@@

%-------------------------------
\subsection{Controller synthesis via state feedback}
\label{ssec:statefeedbackcontrol}
%-------------------------------

@@

%\todo{[Here discuss Jury criterium for DISCRETE-TIME models - a simplified version of the first part of what is now in Section \ref{sec:cof_verification} (with no quantisation noises yet).]}

@@

%%%-------------------------------
%%\subsection{Digital implementation of feedback controller}
%%\label{ssec:quantizationerror}
%%%-------------------------------
%%
%%\todo{HERE add a figure similar to Fig. \label{fig:observersystem} to explain digital features of controller.} 
%%
%%\todo{[The remainder of the section needs to be updated. ]}
%%The output and measurable of the plant is being sampled by an ADC, and since the full state of the
%%plant cannot be witnessed directly, the controller software includes an observer that 
%%estimates the state of the plant (see appendix for details on how such an observer
%%is constructed). In this work, we are assuming that all the states measurements are available for the state feedback purpose, and the state observer design may be disregarded.  
%%\todo{[Unclear what the remainder is for: ]}
%%We further presume that there are small tolerances in the system 
%%described in the following form
%%%
%%\begin{align*}
%%\mat{A} \in \mathcal{G}_A=\mat{S}\mat{J}\mat{S}^{-1} : \mat{J} \in \mathcal{G}_J
%%\end{align*}
%%\commentib{What is S?}
%%
%%Where $\mathcal{G}_J$ is a collection of eigenvalue ranges that define the frequency response of the system.
%%We presume that the internal model of the plant is an ARMA model, modified by a
%%state transform that has no tolerances. Tolerances in the eigenvectors will manifest themselves
%%as tolerances in the canonical forms outside of the row/column of interests and complicate the 
%%evaluation. Whilst it is possible to address these with this methodology, we will focus on the case
%%above for simplicity.

% \todo{HERE add a figure similar to Fig. \label{fig:observersystem} to explain digital features of controller.}

% \todo{[The remainder of the section needs to be updated. ]}
% The output and measurable of the plant is being sampled by an ADC, and since the full state of the
% plant cannot be witnessed directly, the controller software includes an observer that 
% estimates the state of the plant (see appendix for details on how such an observer
% is constructed). In this work, we are assuming that all the states measurements are available for the state feedback purpose, and the state observer design may be disregarded.  
% \todo{[Unclear what the remainder is for: ]}

@@

We require that the closed-loop system meets safety specifications. 
A safety specification establishes bounds for the states, such that 
the controller $K$ must maintain the states within a given lower and upper 
bound for any initial state.

The safety property represented by $\phi_{safety}$ is encoded as follows:
%
\begin{equation}
\label{eq:safetyliteral}
\phi_{safety}\iff \forall k\geq 0, \bigwedge_{i=1}^{n}{\underline{x_{i}} \leq x_{i,k} \leq \overline{x_{i}}},
\end{equation}
%
%\todo{We consider $x_{i}^{-}$ and $x_{i}^{+}$ to be -1 and 1, respectively.}
where $\underline{x_{i}}$ and $\overline{x_{i}}$ are the lower and upper bound 
of the $i$-th state $x_{i}$ at the $k$-th instant, respectively.
This means that the states will always be within the $n$-dimensional 
hyper-box, which is typically defined by the safety specification.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Our synthesis problem}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
The synthesis problem we are trying to solve is the following:
find a digital controller $K$ (see Eq.~\ref{eq:controlaction})
that makes the closed-loop system safe for 
initial state $x_0$, reference signal $r_k$ and input $u_k$
as defined in Sec.~\ref{ssec:safety}. 
We consider non-deterministic initial states within a specified range
and the reference signal to be 0 (\todo{refer to wherever this is explained initially}).
Then, when mapping back to the notation used for describing the general architecture 
of the program synthesizer, the controller $K$ denotes $P$ and 
$(x_0, u_k)$ represents $x$. 
While the controller's precision $\langle I_C,F_C\rangle$ is given, 
we can vary the plant's precision $\langle I_p,F_p\rangle$ such that 
$\mathbb{R}\langle I_p,F_p \rangle \supseteq \mathbb{R}\langle I_c,F_c \rangle$.

@@

\todo{Revise: }
The first aproach finds a controller for an individual initial state and input with a bounded time horizon and generalizes to all states, inputs and time horizons (with a final verification loop that verifies the continuous system behaviour against the previously verified discrete system bisimulation).
The second approach finds a controller for a continuous initial state-space and input sets over an abstraction of the continuous dynamics which conforms to witness proofs at specific times k, and generalizes the solution for all times by refining the number of witnesses.

@@

\todo{Can we come up with some sort of figure that makes it easier to visualise
the generalisation step in each approach?}

@@

\todo{Move as much as possible from Sec~\ref{sec:cof_verification}
to the appendix (e.g. discussion on abstract acceleration). 
Merge the rest in Sec~\ref{sec:cof_synth}.}

@@

 \todo{discuss the computation
  of the completeness threshold}

5. The last {\sc verify} stage, {\sc sampling}, 
ensures that the current sampling rate is large enough by using an 
external checker (\todo{add more details here once we have them}).

@@

LTI models in the form: 
%
\begin{align}
\label{eq:ode}
\dot{\vec{x}}(t)&=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)\ \ :\ \ \vec{x} \in \mathbb{R}^{n}, \vec{u} \in \mathbb{R}^m, \mat{A} \in \mathbb{R}^{n \times n},\mat{B} \in \mathbb{R}^{n \times m}\\
\vec{y}(t)&=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t)\ \ :\ \ \vec{y} \in \mathbb{R}^{o}, \mat{C} \in \mathbb{R}^{o \times n}, \mat{D}  \in \mathbb{R}^{o \times m}\nonumber
\end{align}
\noindent where $\dot{\vec{x}}(t)$ describes the state evolution equation; 
$\vec{y}(t)$ represents an instantaneous output equation; and $\mat{A}$, $\mat{B}$, $\mat{C}$, and $\mat{D}$ are matrices that fully specify 
a continuous plant. Eq.~\eqref{eq:ode} is soundly discretized 
(as shown in the appendix) into
%
\begin{align}
\label{eq:plant}
\vec{x}_{k+1}&=\mat{A}_d\vec{x}_k+\mat{B}_d\vec{u}_k\\
\vec{y}_k&=\mat{C}_d\vec{x}_k+\mat{D}_d\vec{u}_k .\nonumber
\end{align}
%

>>>>

\usepackage{todonotes}
\usetikzlibrary{positioning, automata, shapes.arrows, calc, shapes, arrows}
\usetikzlibrary{patterns}

\newcommand{\addtodo}[1]{{\color{red} TODO: #1}}
\newcommand{\commentib}[1]{{\color{blue} [IB: #1]}}
\newcommand{\mat}[1]{\boldsymbol{#1}}

@@

\begin{abstract}
This work presents a sound and automated approach to synthesise 
digital feedback control architectures for physical plants represented 
as linear, time invariant models. Models are encompassed by dynamical 
equations with inputs, evolving over a continuous state space describing 
the evolution of the physical variables of the system.  The approach is divided 
in two main parts, brought together by a counter-example guided inductive 
synthesis (CEGIS) loop, as follows. First, we devise a static feedback controller 
that stabilises the system; then, we verify its potential safety by means of either 
of the following approaches: forward reachability analysis via unfolding of the dynamics, 
or invariant set generation via abstract acceleration;  
further, if the above step fails, we loop back and invoke a CEGIS call 
to seek alternative stabilising feedbacks, benefitting from the obtained counter-examples 
to safety. 
\textcolor{red}{[Mention GA \todo{GA? Genetic Algorithms? We don't use GA?} and Abstract Acceleration here?]}

The approach is proven sound by accounting for errors due to the digitalisation of signals that are manipulated by the controller block.  
The proposed synthesis approaches and a developed tool are evaluated considering 
a set of intricate physical plant models taken from the digital control literature.
\keywords{
State-space dynamical models of physical systems; 
Digital controllers; 
Analogue-to-digital converters; 
Time sampling; 
Quantisation; 
Fixed-point arithmetics; 
CEGIS; 
safety requirements. 
}
\end{abstract}

@@

In this work, we propose the synthesis of \emph{safe} control algorithms for
LTI models taking into consideration the continuous domain of the plant, 
alongside the discrete domain of the controller and the hybrid elements that relate them.  
%% In this paper, we synthesize controllers for hybrid
%% closed-loop systems, i.e., software-implemented embedded controllers along
%% with a model of their physical environment (the plant), that are
%% \emph{safe}.
Due to the complexity of such systems, in this particular work we focus
on linear models \addtodo{[clarify next] with known configurations} and perform parametric synthesis
of safe digital controllers.

@@

\addtodo{Add contributions (Cristina)}

@@

%\addtodo{[Move material from \ref{sec:rw} here]}

@@

%
\missingfigure{state feedback controller block diagram}
%-------------------------------
\subsection{Controller synthesis via state feedback}
\label{ssec:statefeedbackcontrol}
%-------------------------------

@@

%\addtodo{[Here discuss Jury criterium for DISCRETE-TIME models - a simplified version of the first part of what is now in Section \ref{sec:cof_verification} (with no quantisation noises yet).]}

@@

We also require that the closed-loop system meets safety specifications. 
A safety specification establishes bounds for the states, such that 
the controller $K$ must maintain the states within a given lower and upper 
bound for any initial state. Note that a stable system is not necessarily a
safe system, the state values may leave the safe states during the stabilization.

The safety property represented by $\phi_{safety}$ is encoded as follows:
%
\begin{equation}
\label{eq:safetyliteral}
\phi_{safety}\iff \forall k\geq 0, \bigwedge_{i=1}^{n}{\underline{x_{i}} \leq x_{i,k} \leq \overline{x_{i}}},
\end{equation}
%
%\addtodo{We consider $x_{i}^{-}$ and $x_{i}^{+}$ to be -1 and 1, respectively.}
where $\underline{x_{i}}$ and $\overline{x_{i}}$ are the lower and upper bound 
of the $i$-th state $x_{i}$ at the $k$-th instant, respectively.
This means that the states will always be within the $n$-dimensional 
hyper-box, which is typically defined by the safety specification.

%%%-------------------------------
%%\subsection{Digital implementation of feedback controller}
%%\label{ssec:quantizationerror}
%%%-------------------------------
%%
%%\addtodo{HERE add a figure similar to Fig. \label{fig:observersystem} to explain digital features of controller.} 
%%
%%\addtodo{[The remainder of the section needs to be updated. ]}
%%The output and measurable of the plant is being sampled by an ADC, and since the full state of the
%%plant cannot be witnessed directly, the controller software includes an observer that 
%%estimates the state of the plant (see appendix for details on how such an observer
%%is constructed). In this work, we are assuming that all the states measurements are available for the state feedback purpose, and the state observer design may be disregarded.  
%%\addtodo{[Unclear what the remainder is for: ]}
%%We further presume that there are small tolerances in the system 
%%described in the following form
%%%
%%\begin{align*}
%%\mat{A} \in \mathcal{G}_A=\mat{S}\mat{J}\mat{S}^{-1} : \mat{J} \in \mathcal{G}_J
%%\end{align*}
%%\commentib{What is S?}
%%
%%Where $\mathcal{G}_J$ is a collection of eigenvalue ranges that define the frequency response of the system.
%%We presume that the internal model of the plant is an ARMA model, modified by a
%%state transform that has no tolerances. Tolerances in the eigenvectors will manifest themselves
%%as tolerances in the canonical forms outside of the row/column of interests and complicate the 
%%evaluation. Whilst it is possible to address these with this methodology, we will focus on the case
%%above for simplicity.

% \addtodo{HERE add a figure similar to Fig. \label{fig:observersystem} to explain digital features of controller.}

% \addtodo{[The remainder of the section needs to be updated. ]}
% The output and measurable of the plant is being sampled by an ADC, and since the full state of the
% plant cannot be witnessed directly, the controller software includes an observer that 
% estimates the state of the plant (see appendix for details on how such an observer
% is constructed). In this work, we are assuming that all the states measurements are available for the state feedback purpose, and the state observer design may be disregarded.  
% \addtodo{[Unclear what the remainder is for: ]}

@@

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Our synthesis problem}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
The synthesis problem we are trying to solve is the following:
find a digital controller $K$ (see Eq.~\ref{eq:controlaction})
that makes the closed-loop system safe for 
initial state $x_0$, reference signal $r_k$ and input $u_k$
as defined in Sec.~\ref{ssec:safety}. 
We consider non-deterministic initial states within a specified range
and the reference signal to be 0 (\addtodo{refer to wherever this is explained initially}).
Then, when mapping back to the notation used for describing the general architecture 
of the program synthesizer, the controller $K$ denotes $P$ and 
$(x_0, u_k)$ represents $x$. 
While the controller's precision $\langle I_C,F_C\rangle$ is given, 
we can vary the plant's precision $\langle I_p,F_p\rangle$ such that 
$\mathbb{R}\langle I_p,F_p \rangle \supseteq \mathbb{R}\langle I_c,F_c \rangle$.

@@

\addtodo{Revise: }
The first aproach finds a controller for an individual initial state and input with a bounded time horizon and generalizes to all states, inputs and time horizons (with a final verification loop that verifies the continuous system behaviour against the previously verified discrete system bisimulation).
The second approach finds a controller for a continuous initial state-space and input sets over an abstraction of the continuous dynamics which conforms to witness proofs at specific times k, and generalizes the solution for all times by refining the number of witnesses.

@@

\addtodo{Can we come up with some sort of figure that makes it easier to visualise
the generalisation step in each approach?}

@@

\addtodo{Move as much as possible from Sec~\ref{sec:cof_verification}
to the appendix (e.g. discussion on abstract acceleration). 
Merge the rest in Sec~\ref{sec:cof_synth}.}

@@

 \addtodo{discuss the computation
  of the completeness threshold}

5. The last {\sc verify} stage, {\sc sampling}, 
ensures that the current sampling rate is large enough by using an 
external checker (\addtodo{add more details here once we have them}).

@@

LTI models in the form: 
%
\begin{align}
\dot{\vec{x}}(t)&=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)\ \ :\ \ \vec{x} \in \mathbb{R}^{n}, \vec{u} \in \mathbb{R}^m, \mat{A} \in \mathbb{R}^{n \times n},\mat{B} \in \mathbb{R}^{n \times m}\\
\vec{y}(t)&=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t)\ \ :\ \ \vec{y} \in \mathbb{R}^{o}, \mat{C} \in \mathbb{R}^{o \times n}, \mat{D}  \in \mathbb{R}^{o \times m}\nonumber
\end{align}
\noindent where $\dot{\vec{x}}(t)$ describes the state evolution equation; 
$\vec{y}(t)$ represents an instantaneous output equation; and $\mat{A}$, $\mat{B}$, $\mat{C}$, and $\mat{D}$ are matrices that fully specify 
a continuous plant. Eq.~\eqref{eq:ode} is soundly discretized 
(as shown in the appendix) into
%
\begin{align}
\vec{x}_{k+1}&=\mat{A}_d\vec{x}_k+\mat{B}_d\vec{u}_k\\
\vec{y}_k&=\mat{C}_d\vec{x}_k+\mat{D}_d\vec{u}_k .\nonumber
\end{align}
%
